### 2017-09-25 TF-vldb2018/sec-appendix.tex

\newpage
%\addtolength{\parskip}{0.25cm}
%\newpage

\addtolength{\parskip}{0.25cm}

\vspace{2ex}
\noindent
{\large \textbf{Appendix A: Detailed Algorithms}}
\label{sec-apd-algorithms}

\noindent
{\textbf{1. Procedure \rgraphsim (Section~\ref{subsec-teamF})}}

\vspace{-2ex}
\begin{figure}[h!]
%\vspace{-2ex}
\begin{center}
{\small
\myhrule
\vspace{-2ex}
\mat{0ex}{
\sstab {\sl Input:\/} Pattern graph $P(V_{P}, E_{P})$, ball $\ball([v, r])$.\\
{\sl Output:\/} The perfect subgraph $G_s$ of $\ball{[v,r]}$ \wrt $P$.\\
\sstab \bcc \hspace{1ex} \= \For each $u \in V_{P}$ \Do\\
\icc \>\hspace{2ex}\= $\matchs(u)$ := \{$w | w$ is in $\ball{[v, r]}$ and $l_P(u)\in l(w)$\};\\
\icc \> \While there are changes \Do\\
\icc \>\> \For each edge ($u, u'$) in $E_P$ and each node $w \in \matchs(u)$ \Do\\
\icc \>\>\hspace{2ex}\= \If there is no edge ($w, w'$) in $\ball{[v, r]}$ with $w' \in \matchs(u')$ \Then\\
\icc \>\>\>\hspace{2ex}\= $\matchs(u)$ := $\matchs(u) \setminus \{w\}$;\\
\icc \> \If there exists $|\matchs(u)|\not\in f_{P}(u)$ \Then \Return $\emptyset$;\\
\icc \> $\Reps$ := \{$(u,w) | u \in V_{P}, w \in \matchs(u)$\};\\
\icc \> Construct the perfect subgraph $G_s$ \wrt $\Reps$;\\
\icc \> \Return $G_s$;\\
}

\vspace{-5.5ex} \myhrule
}
\end{center}
\vspace{-4ex}
\caption{Procedure \rgraphsim} \label{alg-rsim}
\vspace{-2ex}
\end{figure}

As shown in Fig~\ref{alg-rsim}, given $P$ and ball $\ball{[v, r]}$,
\rgraphsim finds the perfect subgraph of $P$ and $\ball{[v, r]}$.
For each node $u$ in $V_{P}$, it first computes the set $\matchs(u)$ of candidate matches $w$ containing the label of $u$ (lines 1-2),
and then removes nodes from $\matchs(u)$ iteratively (lines 3-6).
A node $w$ is removed from $\matchs(u)$ if there is a adjacent node $u'$ of $u$,
but there exists no adjacent node $w'$ of $w$ such that $w'\in \matchs(u')$.
Finally, it checks whether the capacity bound on each node $u$ is satisfied (line 7).
If so, \rgraphsim constructs the perfect subgraph $G_s$ \wrt the maximum match relation $\Reps$ and returns it (lines 9-10).

\noindent
{\textbf{2. Algorithm \minp (Section~\ref{subsec-batch-opt})}}

\vspace{-2ex}
\begin{figure}[h!]
%\vspace{-2ex}
\begin{center}
{\small
\myhrule \vspace{-2ex}
\mat{0ex}{
\sstab {\sl Input:\/} Pattern graph $P(V_P$, $E_P)$.\\
{\sl Output:\/} A minimum equivalent pattern graph $P_m$ of $P$.\\
\sstab \bcc \hspace{1ex}\= \If $P$ is not a satisfiable pattern graph \Then \Return \bf{null}.\\
\icc \> Compute the maximum match relation $\Reps$ of $P \eps P$;\\
\icc \> Compute equivalent classes of nodes in $P$ based on $\Reps$;\\
\icc \> Create a node for each equivalent class in $P_m$;\\
\icc \> Connect different equivalent classes by necessary edges in $P_m$;\\
\icc \> Compute the intersection of capacity bounds on nodes in each\\
\>\hspace{0.5ex} equivalent class and put on the node in $P_m$;\\
\icc \> \Return $P_m$;\\
}
\vspace{-6ex}

\myhrule
}
\end{center}
\vspace{-4ex}
\caption{Algorithm \minp} \label{alg-minp}
\vspace{-2ex}
\end{figure}

As shown in Fig.~\ref{alg-minp}, given a pattern $P$, \minp returns a minimum equivalent pattern $P_m$ of $P$.
It firstly checks to see whether $P$ is satisfiable (line 1).
If so, it computes the maximum match relation $\Reps$ by treating $P$ as both pattern and data graph via graph simulation (line 2),
and then computes equivalent classes for nodes in $P$ such that $u$ and $w$ are in the same class iff both $(u, w)\in \Reps$ and $(w, u)\in \Reps$(line 3).
It constructs $P_m$ (lines 4-6) by
(a) creating a class node for $P_m$ for each equivalent class,
(b) connecting the class nodes for $P_m$ iff there exist edges in $P$ connecting two pattern nodes which belong to two equivalent classes, and
(c) computing the intersection of capacity bounds on pattern nodes in each equivalent class and put on the corresponding class node in $P_m$.
\looseness=-1

\vspace{-2ex}
\etitle{Correctness \& Complexity}. The correctness of \minp is assured by following.
(1) The maximum match relations returned by $P$ and $P_{m}$ on any data graph $G$ are always same via graph simulation.
(2) The generated capacity bound on each node in $P_{m}$ assures that, the matched nodes of $P_{m}$ in $G$ always satisfy all capacity bounds on the nodes in $P$, which are in the same equivalent class.
Till now, $P\equiv P_{m}$;
(3) For any $P'$ that $P'\equiv P_{m}$, $|P_{m}|\leq|P'|$.
Algorithm \minp is in $O(|P|^2)$ time, by the time complexity of graph simulation.

\noindent
{\textbf{3. Algorithm \kw{PFrag} (Section~\ref{subsec-framework})}}

Algorithm \kw{PFrag} works by connecting the {\em pattern fragmentation} problem to the
{\sc $(k, \nu)$-Balanced Partition} problem.
It is to divide the nodes of a graph into $k$ components such that each component is of size no more than $\nu\cdot\frac{|V|}{k}$, and the number of edges between different components is minimized.
As illustrated before, the $(k,\nu)$-balanced partition problem,
though is not approximable in general, has a number of sophisticated heuristic algorithms~\cite{AndreevR06}.

\vspace{-2ex}
\begin{figure}[h!]
\begin{center}
{\small
\begin{center}
\myhrule \vspace{-2ex}
\mat{0ex}{
\sstab {\sl Input:\/} \= Pattern graph $P(V_{P},E_{P})$, integer $h$.\\
{\sl Output:\/} An $h$-fragmentation \{${P}_{f1}$, \ldots, ${P}_{fh}$, $C$\} of $P$.\\
\sstab \bcc \hspace{0.5ex} \= $k$ := $h$; $\nu$ := $h$; $M_P$ := $|P|$; $M_C$ := 0;\\
\icc \> $({P}^\nu_{f1}, \ldots, {P}^\nu_{fh}, C^\nu)$ := $\kw{BalanceP}(k,\nu)$; \\
\icc \> $M^\nu_P$ := $\kw{max}\{|{P}^\nu_{f1}|, \ldots, |{P}^\nu_{fh}|\}$; $M^\nu_C$ := $|C^\nu|$;\\
\icc\> \While $\kw{max}\{M_P, M_C\} > \kw{max}\{M^\nu_P, M^\nu_C\}$ \Do\\
\icc \>\hspace{2ex}\= ${P}_{f1}$ := ${P}^\nu_{f1}$, \ldots, ${P}_{fh}$ := ${P}^\nu_{fh}$;\\
\icc \>\hspace{2ex}\= $C$ := $C^\nu$; $M_P$ := $M^\nu_P$; $M_C$ := $M^\nu_C$;\\
\icc \>\> \If $M^\nu_P \geq M^\nu_C$ \Then $\nu := \frac{\nu}{2}$ \Else $\nu := \frac{3\nu}{2}$;\\
\icc \>\> $({P}^\nu_{f1}, \ldots, {P}^\nu_{fh}, C)$ := $\kw{BalanceP}(k,\nu)$;\\
\icc \>\> $M^\nu_P$ := $\kw{max}\{|{P}^\nu_{f1}|, \ldots, |{P}^\nu_{fh}|\}$; $M^\nu_C$ := $|C^\nu|$;\\
\icc \> \Return \{${P}_{f1}$, \ldots, ${P}_{fh}$, $C$\};
}
\vspace{-2.5ex} \myhrule
\end{center}
}
\end{center}
\vspace{-4ex}
\caption{Algorithm \kw{PFrag} \label{alg-pfrag}}
\vspace{-2ex}
\end{figure}

As shown in Fig.~\ref{alg-pfrag}, given $P$ and integer $h$,
\kw{PFrag} finds an $h$-fragmentation for $P$ by recursively invoking algorithm $\kw{BalanceP}(k,\nu)$ for the {\sc $(k,\nu)$-Balanced partition} problem ($\nu\geq 1$) with different $\nu$,
such that the final returned $h$-fragmentation strikes a balance between the size of each fragment ${P}_{fi}$ and the size of the cut $C$.
More specifically, \kw{PFrag} maintains $M_P$ and $M_C$ as the size of the largest fragment ${P}_{fi}$ and the size of the cut respectively.
Initially, it sets $M_P$ to $|P|$ and $M_C$ to 0 (line~1).
It then invokes \kw{BalanceP} with both $k$ and $\nu$ being $h$, \ie has no constraints on the size of fragments of $P$ (line~2).
After that, it iteratively checks whether the generated $h$-fragmentation can be improved (line~4) by adjusting $\nu$ in a {\em binary search style} (lines~4-9).
If the current size $M^\nu_P$ of the largest fragment is no smaller than the size $M^\nu_C$ of the cut,
it invokes \kw{BalanceP} with $h$ and $\frac{\nu}{2}$, or with $h$ and $\frac{3\nu}{2}$ otherwise (line~7).
It returns the $h$-fragmentation if it cannot be improved anymore (line~10).

\etitle{Correctness \& Complexity}.
The correctness is obvious as \kw{PFrag} always returns an $h$-fragmentation.
Algorithm~\kw{PFrag} runs in $O(\log h\cdot t_{\kw{BalanceP}})$, where $t_{\kw{BalanceP}}$ is the complexity of the algorithm for the {\sc $(k,\nu)$-Balanced Partition} problem. Indeed, \kw{PFrag} calls at most $\log h$ times \kw{BalanceP}, as $\nu\geq 1$.
\looseness=-1

\vspace{1ex}
\noindent
{\large \textbf{Appendix B: Detailed Proofs}}
\label{sec-apd-proofs}

\noindent
{\textbf{1. Proof of Proposition~\ref{thm-inc-grouprec-pat}}}

Incremental complexity is defined in terms of LP (locally persistent) graph algorithms \cite{Reps96}.
We also adopt the notion to prove unboundedness of graph algorithms for \dynteamF.

\vspace{-1.5ex}
The proofs below strictly follows the one in \cite{Reps96}.

\vspace{-1.5ex}
\begin{figure}[h!]
\label{fig-inc-complexity-data}
\begin{center}
\includegraphics[scale=0.31]{./fig/inc-complexity-proof-data.eps}
\end{center}
\vspace{-3ex}
\caption{Unboundedness of unit data update}
\vspace{-2ex}
\end{figure}

\etitle{(I) \dynteamF\, is unbounded for unit data update}.
%Following \cite{Reps96, FanWW13-tods},
Consider the following pattern and data graph, and unit data updates.
Let data graph $G$ consist of two chains ($a_1$, $b_1$, $c_1$, \ldots, $a_n$, $b_n$, $c_n$) and ($a_{n+1}$, $b_{n+1}$, $c_{n+1}$, \ldots, $a_{2n}$, $b_{2n}$, $c_{2n}$) where $a_i$, $b_i$ and $c_i$ have labels $A$, $B$ and $C$ respectively.
Let pattern graph be a triangle with nodes $a$, $b$ and $c$ with labels $A$, $B$ and $C$ respectively.
Consider two unit edge insertions $\delta_1$ = $(c_n$, $a_{n+1})^+$ and $\delta_2$ = $(c_{2n}$, $a_1)^+$,
and set $k = 1$ and $r$ = $6n$.
Let $H_1$ and $H_2$ denote the graphs $G\oplus \delta_1$ and $G\oplus \delta_2$, respectively.
Obviously, $L_{k}(P, G) = L_{k}(P, H_1) = L_{k}(P, H_2) = \emptyset$,
while $L_{k}(P, H_1\oplus \delta_2) \neq \emptyset$.
Assume that there exists a locally persistent incremental algorithm ${\cal A}$ for \dynteamF.
Let $Trace(G', \delta')$ denote the sequence of steps executed by ${\cal A}$ in processing some update $\delta'$ to some graph $G'$.
Now consider the following two instances: the application of update $\delta_2$ to $G$ and the application of update $\delta_2$ to graph $H_1$.
Obviously, the update process must behave differently in these two cases, and $Trace(G, \delta_2)$ must be different from $Trace(H_1, \delta_2)$
(because many nodes of $H_1\oplus \delta_2$ are affected, while no node in $G\oplus \delta_2$ is affected).
Since a locally persistent algorithm makes no use of global storage, this can happen only both $Trace(G, \delta_2)$ and $Trace(H_1, \delta_2)$ include a visit to some node $w$ that contains different information in the graphs $G$ and $H_1$.
However, $H_1$ was obtained from $G$ by applying update $\delta_1$.
Hence the information at node $w$ must have been changed during the updating of applying $\delta_1$ to $G$.
Therefore, $Trace(G, \delta_1)$ must also contain a visit to node $w$.
As a characteristic of locally persistent algorithms is that if a node $w$ is visited during the updating of applying change $\delta'$ to graph $G'$,
then every node on some path in $G'$ from a modified node of $\delta'$ to $w$ must have been visited.
Therefore, $Trace(G, \delta_1)$ and $Trace(G, \delta_2)$ both contain a visit to $w$,
from the nodes in $\delta_1$ and $\delta_2$, respectively.
Thus, $Trace(G, \delta_1)$ and $Trace(G, \delta_2)$ include visits to every node on the path from $c_n$ or $a_{n+1}$ to $c_{2n}$ or $a_1$ respectively.
Hence, the time taken for processing update $\delta_1$ to $G$ plus the time taken for processing update $\delta_2$ to $G$ must be no smaller than the distance between $c_n$ or $a_{n+1}$ to $c_{2n}$ or $a_1$, \ie $n$, which is not a constant.
However, $|\kw{AFF}|$ in both cases are 1, such that the complexity of the incremental algorithm ${\cal A}$ cannot be measured by a function of $|\kw{AFF}|$.
Thus, ${\cal A}$ is not a bounded locally persistent incremental algorithm.
%which contradicts the assumption.

\vspace{-1.5ex}
That is, \dynteamF\, is unbounded even for $k=1$ and unit data update.
\looseness=-1

\vspace{-1ex}
\begin{figure}[h!]
\label{fig-inc-complexity-pattern}
\begin{center}
\includegraphics[scale=0.31]{./fig/inc-complexity-proof-pattern.eps}
\end{center}
\vspace{-3ex}
\caption{Unboundedness of unit pattern update}
\vspace{-1.5ex}
\end{figure}

\vspace{-1ex}
\etitle{(II) \dynteamF\, is unbounded for unit pattern update}.
Consider the following pattern and data graph, and unit pattern updates.
Let data graph $G$ be a cycle ($a'_1$, $b'_1$, $c'_1$, $d'_1$, \ldots, $a'_m$, $b'_m$, $c'_m$, $d'_m$, $a'_1$),
where $a'_i$, $b'_i$, $c'_i$ and $d'_i$ have labels $A$, $B$, $C$ and $D$ respectively.
Let pattern graph be a cycle ($a_1$, $b_1$, $c_1$, $d_1$, \ldots, $a_n$, $b_n$, $c_n$, $d_n$, $a_{n+1}$, $b_{n+1}$, $c_{n+1}$, $d_{n+1}$, \ldots, $a_{2n}$, $b_{2n}$, $c_{2n}$, $d_{2n}$, $a_1$) and two extra edges ($a_1$, $c_1$) and ($b_{n+1}$, $d_{n+1}$),
where $a_i$, $b_i$, $c_i$ and $d_i$ have labels $A$, $B$, $C$ and $D$ respectively.
Consider two unit edge deletions $\delta_1$ = $(a_1$, $c_1)^-$ and $\delta_2$ = $(b_{n+1}$, $d_{n+1})^-$,
and set $k = 1$ and $r$ = $4m$.
Let $H_1$ and $H_2$ denote the graphs $P\oplus \delta_1$ and $P\oplus \delta_2$, respectively.
Obviously, $L_{k}(P, G) = L_{k}(H_1, G) = L_{k}(H_2, G) = \emptyset$,
while $L_{k}(H_1\oplus \delta_2, G) \neq \emptyset$.
Assume there exists a locally persistent incremental algorithm ${\cal A}$ for \dynteamF.
Let $Trace(P', \delta')$ denote the sequence of steps executed by ${\cal A}$ in processing some update $\delta'$ to some pattern $P'$.
Now consider two instances: the application of update $\delta_2$ to $P$ and the application of update $\delta_2$ to $H_1$.
Obviously, the update process must behave differently in these two cases, and $Trace(P, \delta_2)$ must be different from $Trace(H_1, \delta_2)$
(because many nodes in $G$ for $H_1\oplus \delta_2$ are affected, while no node in $G$ for $P\oplus \delta_2$ is affected).
Since a locally persistent algorithm makes no use of global storage,
this can happen only both $Trace(P, \delta_2)$ and $Trace(H_1, \delta_2)$ include a visit to some node $w$ that contains different information in $P$ and $H_1$.
However, $H_1$ was obtained from $P$ by applying $\delta_1$.
Hence the information at node $w$ must have been changed during the updating of applying $\delta_1$ to $P$.
Therefore, $Trace(P, \delta_1)$ must also contain a visit to node $w$.
%As a characteristic of locally persistent algorithms is that if a node $w$ is visited during the updating of applying update $\delta'$ to pattern $P'$,
%then every node in some path in $P'$ from a modified node of $\delta'$ to $w$ must have been visited.
According to the characteristics of locally persistent algorithms as illustrated in the data updates case,
$Trace(P, \delta_1)$ and $Trace(P, \delta_2)$ both contain a visit to $w$,
from the nodes in $\delta_1$ and $\delta_2$, respectively.
Thus, $Trace(P, \delta_1)$ and $Trace(P, \delta_2)$ include visits to every node on the path from $a_1$ or $c_1$ to $b_{n+1}$ or $d_{n+1}$ respectively.
Hence, the time taken for processing $\delta_1$ to $P$ plus the time taken for processing $\delta_2$ to $P$ must be no smaller than the distance between $a_1$ or $c_1$ to $b_{n+1}$ or $d_{n+1}$, \ie $4n$, which is not a constant.
However, $|\kw{AFF}|$ in both cases are 1, such that the complexity of algorithm ${\cal A}$ cannot be measured by a function of $|\kw{AFF}|$.
Thus, ${\cal A}$ is not a bounded locally persistent incremental algorithm.
%which contradicts the assumption.
That is, \dynteamF\, is unbounded even for $k=1$ and unit pattern update.

\vspace{-1.5ex}
(1) and (2) together prove that \dynteamF\, is unbounded even for $k=1$ and unit pattern or data update.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tb!]
%\begin{widepage}
\vspace{-2ex}
\begin{center}
%\hspace{-5.5ex}
\subfigure[{\scriptsize Varying $|V_{P}|$ (\youtube)}]{\label{fig-exp-semantic-youtube-diameter}
\includegraphics[scale=0.37]{./fig/youtube-semantic-diameter.eps}}
\hspace{0.2ex}
\subfigure[{\scriptsize Varying $|V_{P}|$ (\youtube)}]{\label{fig-exp-semantic-youtube-density}
\includegraphics[scale=0.37]{./fig/youtube-semantic-density.eps}}
\hspace{0.2ex}
\subfigure[{\scriptsize Varying $|V_{P}|$ (\youtube)}]{\label{fig-exp-semantic-youtube-capacity}
\includegraphics[scale=0.37]{./fig/youtube-semantic-capacity.eps}}
\hspace{0.2ex}
\subfigure[{\scriptsize Varying $|V_{P}|$ (\youtube)}]{\label{fig-exp-semantic-youtube-link}
\includegraphics[scale=0.37]{./fig/youtube-semantic-link.eps}}
\vspace{-2.5ex}

%\hspace{-5.5ex}
\subfigure[{\scriptsize Varying $k$ (\youtube)}]{\label{fig-exp-semantic-youtube-diameter-varyk}
\includegraphics[scale=0.37]{./fig/youtube-semantic-diameter-varyk.eps}}
\hspace{0.2ex}
\subfigure[{\scriptsize Varying $k$ (\youtube)}]{\label{fig-exp-semantic-youtube-density-varyk}
\includegraphics[scale=0.37]{./fig/youtube-semantic-density-varyk.eps}}
\hspace{0.2ex}
\subfigure[{\scriptsize Varying $k$ (\youtube)}]{\label{fig-exp-semantic-youtube-capacity-varyk}
\includegraphics[scale=0.37]{./fig/youtube-semantic-capacity-varyk.eps}}
\hspace{0.2ex}
\subfigure[{\scriptsize Varying $k$ (\youtube)}]{\label{fig-exp-semantic-youtube-link-varyk}
\includegraphics[scale=0.37]{./fig/youtube-semantic-link-varyk.eps}}
\vspace{-2.0ex}

\end{center}
\vspace{-2.0ex}
\caption{Performance evaluation of \optgrouprec for top-$k$ team formation problem}
\label{exp-semantic-effectiveness-youtube}
%\end{widepage}
%\vspace{-0.5ex}
\end{figure*}

\begin{figure*}[tb!]
%\begin{widepage}
\vspace{-1.5ex}
\begin{center}
%\hspace{-5.5ex}
\subfigure[{\scriptsize Varying $|\Delta P|$ (deletions)}]{\label{fig-exp-patinc-del-youtube}
\includegraphics[scale=0.395]{./fig/youtube-pattern-deletion.eps}}
\hspace{-1.7ex}
\subfigure[{\scriptsize Varying $|\Delta P|$ (insertions)}]{\label{fig-exp-patinc-ins-youtube}
\includegraphics[scale=0.395]{./fig/youtube-pattern-insertion.eps}}
\hspace{-1.7ex}
\subfigure[{\scriptsize Varying $|\Delta P|$ (capacity)}]{\label{fig-exp-patinc-cap-youtube}
\includegraphics[scale=0.395]{./fig/youtube-pattern-capchange.eps}}
\hspace{-1.7ex}
\subfigure[{\scriptsize Varying $|\Delta P|$ (hybrid updates)}]{\label{fig-exp-patinc-hyb-youtube}
\includegraphics[scale=0.395]{./fig/youtube-pattern-hybrid.eps}}
\vspace{-2.5ex}

%\hspace{-5.5ex}
\subfigure[{\scriptsize Varying $|\Delta G|$ (deletions)}]{\label{fig-exp-datainc-del-youtube}
\includegraphics[scale=0.395]{./fig/youtube-data-deletion.eps}}
\hspace{-1.7ex}
\subfigure[{\scriptsize Varying $|\Delta G|$ (insertions)}]{\label{fig-exp-datainc-ins-youtube}
\includegraphics[scale=0.395]{./fig/youtube-data-insertion.eps}}
\hspace{-1.7ex}
\subfigure[{\scriptsize Varying $|\Delta G|$ (hybrid updates)}]{\label{fig-exp-datainc-hyb-youtube}
\includegraphics[scale=0.395]{./fig/youtube-data-hybrid.eps}}
\hspace{-1.7ex}
\subfigure[{\scriptsize Vary $(|\Delta P|,|\Delta G|)$ (simultaneous)}\hspace{-4.5ex}]{\label{fig-exp-hyb-patdata-youtube}
\includegraphics[scale=0.395]{./fig/youtube-pattern-data.eps}}
\vspace{-2.5ex}

%\hspace{-5.5ex}
\subfigure[{\scriptsize Varying $|\Delta P|$ in update sets}]{\label{fig-exp-patinc-multi-youtube}
\includegraphics[scale=0.395]{./fig/youtube-pattern-multi.eps}}
%\hspace{-1.7ex}
\subfigure[{\scriptsize Varying $|\Delta G|$ in update sets}]{\label{fig-exp-datainc-multi-youtube}
\includegraphics[scale=0.395]{./fig/youtube-data-multi.eps}}
%\hspace{-1.7ex}
\subfigure[{\scriptsize Varying $(|\Delta P|,|\Delta G|)$ in update sets}\hspace{-5.5ex}]{\label{fig-exp-hyb-datapat-multi-youtube}
\includegraphics[scale=0.395]{./fig/youtube-pattern-data-multi.eps}}
%\hfill
\vspace{-2.5ex}

\end{center}
\vspace{-2.0ex}
\caption{Performance evaluation of \inc for dynamic top-$k$ team formation problem (\youtube)}
\label{exp-inc-youtube}
%\end{widepage}
\vspace{-4ex}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\noindent
{\textbf{2. Proof of Proposition~\ref{prop-fragmentation}}}

The decision version of the {\em pattern fragmentation} problem, denoted by $\kw{dOFGP}(P, h, r_1, r_2)$,
is to decide whether there exists a fragmentation \{${P}_{f1}$, \ldots, $P_{fh}$, $C$\} such that,
(a) $\max_i |P_{fi}| \leq r_1\cdot\frac{|P|}{h}$ and (b) $|C|\leq r_2|P|$.

\vspace{-2ex}
\etitle{Upper bound.} We show the \NP upper bound by providing an \NP algorithm to determine whether there exists an $h$-fragmentation of $P$.\,Given $P$, the algorithm works as follows.
\looseness=-1

\vspace{-1.8ex}
\sstab{(1)} Guess an $h$-fragmentation ${\cal P}_h$ of $P$.

\vspace{-1.8ex}
\sstab{(2)} Check whether it satisfies restrictions of $r_1$ and $r_2$ (conditions (a) and (b) in the definition of $\kw{dOFGP}$).
If so, return yes, otherwise go to the first step and guess another instance.

\vspace{-1.5ex}
The algorithm is in \NP since step~(2) can be checked in \PTIME (liner time, indeed).

\vspace{-2.5ex}
\etitle{Lower bound.}
We next show that it is \NP-hard to determine whether there exists an $h$-fragmentation and remains \NP-hard even when $h$ = 2,
by reduction from the {\em minimum cut} problem, which is known \NP-complete\footnote{\small http://www.nada.kth.se/$\sim$viggo/wwwcompendium/node85.html}.

\vspace{-2ex}
An instance of the minimum cut problem is given an undirected graph $H=(V_{H},E_{H})$ and a positive integer $m$,
to find a set $S \subseteq E_{H}$ of edges whose removal leaves two disjoint connected components, with $|S|\leq m$.

\vspace{-2ex}
Given an instance of the minimum cut problem,
we construct an instance of $h$-fragmentation problem.
It can be achieved by taking $H$ as $P$, and setting $h=2$, $r_1=h$ and $r_2 = m/|P|$.
Indeed, the minimum cut problem is a special case of the of $h$-fragmentation problem.

\noindent
{\textbf{3. Proof of Theorem~\ref{thm-compose}}}

We will prove the theorem by induction.
Given pattern $P(V_P,E_P)$ and its fragmentation \{${P}_{f1}$, \ldots, ${P}_{fh}$, $C$\},
we use $P^{C}(V_{P^{C}},E_{P^{C}})$ to denote the pattern with $V_{P^{C}}$=$V_P$ and $E_{P^{C}}$=$E_P/C$.
%we use $P^{C}(V_P,E_P/C)$ to denote the pattern with the same node set $V_P$ with $P$, and edges excluding the cut $C$ from $E_P$.
Graph simulation is an iterative process to remove unmatched nodes from the candidate nodes,
as illustrated in Fig.~\ref{alg-rsim} (lines~1-6).
We utilize $M_{C}^{k}$ (resp. $M_{i}^{k}$, $M^{k}$) to denote the match relation for $P^{C}$ (resp. $P_{fi}$, $P$) in $\ball{}$ in the $kth$ iteration.
By the definition of graph simulation, we have $M_{C}^{k}$ = $\bigcup_{i=1}^{h}M_{i}^{k}$.
To prove $M\subseteq\bigcup_{i=1}^{h}M_{i}$,
we next prove $M^{k}\subseteq M_{C}^{k}$ for each iteration instead.

\vspace{-1.8ex}
\sstab{(1)} For $k=0$, \ie the initialization step of graph simulation algorithm,
the algorithm computes the set of candidate matches for each pattern node with the same label.
As $P$ and $P^{C}$ have exactly the same node set, we have $M^{0}= M_{C}^{0}$.

\vspace{-1.8ex}
\sstab{(2)} For $k=n$ $(n\geq 0)$, if $M^{n}\subseteq M_{C}^{n}$ exists, we prove $M^{n+1}\subseteq M_{C}^{n+1}$ exists in the $(n+1)th$ iteration.
Suppose both $(u,w)\in M^{n}$ and $(u,w)\in M_{C}^{n}$ exist,
and in the $(n+1)th$ iteration, $(u,w)$ is removed from $M_{C}^{n}$ if there is a adjacent node $u'$ of $u$ in $P_{C}$,
but there exists no adjacent node $w'$ of $w$ in $G$ such that $(u',w')\in M_{C}^{n}$.
Therefore, $(u,w)$ must be removed from $M^{n}$ as $E_{P^{C}} \subseteq E_{P}$,
that is, the edge $(u,u')$ in $E_{P^{C}}$ must belong to $E_{P}$.
Thus, we have $M^{n+1}\subseteq M_{C}^{n+1}$.

\vspace{-1.5ex}
By (1) and (2), we have proven that $M\subseteq\bigcup_{i=1}^{h}M_{i}$.

\vspace{1ex}
\noindent
{\large \textbf{Appendix C: Extra Experiments}}
\label{sec-apd-exp}

\vspace{-0.5ex}
The experimental results on a real-life dataset \youtube are reported here.

{\em \youtube}~\cite{youtubeWeb} contains 2.03M video nodes and 12.2M edges, which represent recommendations between two videos. We used the undirected version, and generated 400 labels based on the built-in categories and ages of videos.

\vspace{-1.8ex}
\stitle{Exp-1: Performance of \optgrouprec}. We firstly evaluated the performance of \optgrouprec vs. \mindia, \minsumdis, \denalk on \youtube \wrt three measurements.
The results are reported in Fig.~\ref{fig-exp-semantic-youtube-diameter} to~\ref{fig-exp-semantic-youtube-link-varyk}.
We find \optgrouprec strikes a balance at capturing the practical requirements.

\vspace{-1.8ex}
\stitle{Exp-2: Efficiency of \inc for one set of updates}. Varying the amount of updates in one update set from 4.5\% to 45\% for pattern updates, 2.5\% to 25\% for data updates, and (4.5\%, 2.5\%) to (31.5\%, 17.5\%) for simultaneous updates, the results are reported in Fig.~\ref{fig-exp-patinc-del-youtube} to~\ref{fig-exp-hyb-patdata-youtube}.
We find \inc outperforms \optgrouprec when $\Delta P$, $\Delta G$ and $(\Delta P$, $\Delta G)$ are no more than 31.5\%, 20\% and (27\%, 15\%).
%for hybrid pattern, hybrid data, and simultaneous pattern and data updates.

\vspace{-1.8ex}
\stitle{Exp-3: Efficiency of \inc for continuous sets of updates}. We generated 5 sets of updates,
varying the amount of updates from 4.5\% to 27\% for pattern updates, 2.5\% to 15\% for data updates,
and (4.5\%, 2.5\%) to (22.5\%, 12.5\%) for simultaneous pattern and data updates.
We evaluated the average time took by \inc to process these sets of updates one by one.
The results are reported in Fig.~\ref{fig-exp-patinc-multi-youtube} to~\ref{fig-exp-hyb-datapat-multi-youtube}.
We find \inc outperforms \optgrouprec when changes are no more than 27\%, 10\% and (13.5\%, 7.5\%) for continuous pattern, data and simultaneous updates respectively.

\vspace{-1.8ex}
\stitle{Exp-4: Physical storage of auxiliary structures}. It takes 58MB extra space to store all auxiliary structures utilized by \inc on \youtube, which need 114MB space to store itself,
\ie 50.8\% compared with the original dataset.