### 2017-09-25 TF-vldb2018/sec-conclusion.tex

\vspace{-2ex}
\section{Conclusion}
\label{sec-conclusion}

We have introduced a graph pattern matching approach for (dynamic) top-$k$ team formation problem.
We have proposed team simulation,
%based on which we have developed an efficient batch algorithm for top-k team formation.
and also  developed a unified incremental framework to handle continuous pattern and data updates, separately and simultaneously.
We have experimentally verified the effectiveness and efficiency of the batch and incremental algorithms.
\looseness=-1

A couple of topics are targeted for future work.  First, an interesting topic is to develop distributed algorithms for top-$k$ team formation.
Second, the study of dynamic algorithms for query updates is in its infancy,
and hence, an important topic is to develop such algorithms for various problems.