### 2017-09-28 Gruba/08-conclu.tex

\section{\textcolor{red}{Conclusions}}
\label{sec:conclu}

%zhu begin

In this paper, we studied to model reposting behavior of group users in microblog. We first extracted users's features including basic features, behavior features and interests features. Meanwhile, we proposed a novel method to extract user interests from the user generated texts. By performing users clustering, we divided users into several groups and modeled reposting behavior of each group users respectively. The final experiment shows that our model have strong predictive power. What's more, we developed a demo system to display the profiling for users and groups. As future work, it is interesting to study other methods to improve the accuracy of interests extraction.