### 2017-09-30 Gruba/07-rela.tex
v2.1
v2.1

\section{\textcolor{red}{Related Work}}
\label{sec:rela}