### 2017-09-30 Gruba/08-conclu.tex

\section{\textcolor{red}{Conclusions}}
\label{sec:conclu}