### 2017-10-02 Gruba/08-conclu.tex
v3.0
v3.0

\section{Conclusions}
\label{sec:conclu}

In this work, we have presented \sys{}, a system to model users' \retg{} behavior in social media. 
\sys{} departures from related works by grouping users into clusters, during which features of info, behavior and interests are involved. 
Specially, we have studied interest features from various perspectives, such as long-term/recent interests and explicit/implicit interests. 
Last, we provided the performance evaluation of \sys{} by using real-world datasets and comparing against state of the art approaches, showing its benefits.