### 2017-10-06 Gruba/10-appendix.tex

\newpage
\section{Appendix: Extra Experiments}
\label{sec:appendix}

\subsection{Case Study for Feature Extraction}

\subsection{Case Study for User Clustering}