### 2017-10-13 Gruba/08-conclu.tex

\section{Conclusions}
\label{sec:conclu}

In this work, we have presented \sys{}, a system to model the \retg{} behavior of users in social media.
\sys{} departures from existing work by grouping users into clusters, during which features of basics, behavior and interests are extracted.
Specially, we have studied interest features from various perspectives, such as long-term/recent interests and explicit/implicit interests.
Finally, we have provided a performance evaluation of \sys{} by using real-world datasets to demonstrate the benefits of our system \sys{}.