### 2017-10-13 Gruba/10-appendix.tex

\newpage
\section{Appendix: Extra Experiments}
\label{sec:appendix}

%\subsection{Case Study for Feature Extraction}

\subsection{Case Study for User Clustering}

In this section, we show the results of our demo system for User Clustering. We carefully analyzed the statistical information of four groups obtained from user clustering.

\stab(1) Most users of group one are male (as shown in Sub-Fig. \ref{fig:subfig1:fig11}), and Sub-Fig. \ref{fig:subfig1:fig12} shows that they are mainly in Beijing(北京). Most of them have number of followees similar as followers (as shown in Sub-Fig. \ref{fig:subfig1:fig14}). The most active time of this group is mainly between 10 am and 12 am (as shown in Sub-Fig. \ref{fig:subfig1:fig15} and \ref{fig:subfig1:fig16}).

\stab(2) The users of group two are mostly female (as shown in Sub-Fig. \ref{fig:subfig2:fig21}) and are also mainly in Beijing (北京) (shown as Sub-Fig. \ref{fig:subfig2:fig22}). Similar as group one, most of them have number of followees close to followers (as shown in Sub-Fig. \ref{fig:subfig2:fig24}). The most active time of this group is mainly between 10 pm and 12 pm (as shown in Sub-Fig. \ref{fig:subfig2:fig25} and \ref{fig:subfig2:fig26}).

\stab(3) Most users of group three are male (as shown in Sub-Fig. \ref{fig:subfig3:fig31}). They come from a wide range of provinces (as shown in Sub-Fig. \ref{fig:subfig3:fig32}). Most of them have number of followers far more than followees (as shown in Sub-Fig. \ref{fig:subfig3:fig34}). The most active time of this group is mainly in 10-12 am and 10-12 pm (as shown in Sub-Fig. \ref{fig:subfig3:fig35} and \ref{fig:subfig3:fig36}).

\stab(4) Most users of group four are female (as shown in Sub-Fig. \ref{fig:subfig4:fig41}), and they come from a wide range of provinces just as group three (as shown in Sub-Fig. \ref{fig:subfig4:fig42}). Most of these users have number of followers far more than followees (as shown in Sub-Fig. \ref{fig:subfig4:fig44}). And the most active time of these users is mainly between 10 pm and 12 pm (as shown in Sub-Fig. \ref{fig:subfig4:fig45} and \ref{fig:subfig4:fig46}).

According to Sub-Fig. \ref{fig:subfig1:fig13}, \ref{fig:subfig2:fig23}, \ref{fig:subfig3:fig33} and \ref{fig:subfig4:fig43}, 
the number distribution of microblogs for four groups is similar with each other.
We visualized each group's long-term interest expressed as words cloud (shown in Sub-Fig. \ref{fig:subfig1:fig19}, \ref{fig:subfig2:fig29}, \ref{fig:subfig3:fig39} and \ref{fig:subfig4:fig49}), as we can see, group one and group three have similar interest.

\begin{figure*}
  \centering
  \subfigure{
  \label{fig:subfig1:fig11}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/11.eps}}
  \subfigure{
  \label{fig:subfig1:fig12}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/12.eps}}
  \subfigure{
  \label{fig:subfig1:fig13}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/13.eps}}
  \subfigure{
  \label{fig:subfig1:fig14}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/14.eps}}
  \subfigure{
  \label{fig:subfig1:fig15}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/15.eps}}
  \subfigure{
  \label{fig:subfig1:fig16}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/16.eps}}
  \subfigure{
  \label{fig:subfig1:fig17}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/17.eps}}
  \subfigure{
  \label{fig:subfig1:fig18}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/18.eps}}
  \subfigure{
  \label{fig:subfig1:fig19}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/19.eps}}
  \caption{The Statistics of User Group One}
  \label{fig:subfig1} %% label for entire figure
\end{figure*}

\begin{figure*}
  \centering
  \subfigure{
  \label{fig:subfig2:fig21}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/21.eps}}
  \subfigure{
  \label{fig:subfig2:fig22}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/22.eps}}
  \subfigure{
  \label{fig:subfig2:fig23}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/23.eps}}
  \subfigure{
  \label{fig:subfig2:fig24}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/24.eps}}
  \subfigure{
  \label{fig:subfig2:fig25}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/25.eps}}
  \subfigure{
  \label{fig:subfig2:fig26}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/26.eps}}
  \subfigure{
  \label{fig:subfig2:fig27}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/27.eps}}
  \subfigure{
  \label{fig:subfig2:fig28}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/28.eps}}
  \subfigure{
  \label{fig:subfig2:fig29}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/29.eps}}
  \caption{The Statistics of User Group Two}
  \label{fig:subfig2} %% label for entire figure
\end{figure*}

\begin{figure*}
  \centering
  \subfigure{
  \label{fig:subfig3:fig31}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/31.eps}}
  \subfigure{
  \label{fig:subfig3:fig32}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/32.eps}}
  \subfigure{
  \label{fig:subfig3:fig33}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/33.eps}}
  \subfigure{
  \label{fig:subfig3:fig34}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/34.eps}}
  \subfigure{
  \label{fig:subfig3:fig35}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/35.eps}}
  \subfigure{
  \label{fig:subfig3:fig36}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/36.eps}}
  \subfigure{
  \label{fig:subfig3:fig37}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/37.eps}}
  \subfigure{
  \label{fig:subfig3:fig38}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/38.eps}}
  \subfigure{
  \label{fig:subfig3:fig39}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/39.eps}}
  \caption{The Statistics of User Group Three}
  \label{fig:subfig3} %% label for entire figure
\end{figure*}

\begin{figure*}
  \centering
  \subfigure{
  \label{fig:subfig4:fig41}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/41.eps}}
  \subfigure{
  \label{fig:subfig4:fig42}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/42.eps}}
  \subfigure{
  \label{fig:subfig4:fig43}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/43.eps}}
  \subfigure{
  \label{fig:subfig4:fig44}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/44.eps}}
  \subfigure{
  \label{fig:subfig4:fig45}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/45.eps}}
  \subfigure{
  \label{fig:subfig4:fig46}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/46.eps}}
  \subfigure{
  \label{fig:subfig4:fig47}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/47.eps}}
  \subfigure{
  \label{fig:subfig4:fig48}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/48.eps}}
  \subfigure{
  \label{fig:subfig4:fig49}
      \includegraphics[width=0.23\textwidth]{IMAGE/group-images/49.eps}}
  \caption{The Statistics of User Group Four}
  \label{fig:subfig4} %% label for entire figure
\end{figure*}