### 2017-12-17 DASFAA_2018/Gruba/00-abstr.tex

\begin{abstract}
%Social media applications are emerging, with rapidly growing users and large numbers of \retd{} microblogs every minute.
The variety among massive users makes it difficult to model their \retg{} activities.
Obviously, it is not suitable to cover the overall users by a single model. % \cite{IEEEexample:journals/tkdd/ZhangTLLX15,IEEEexample:conf/wsdm/FengW13,IEEEexample:conf/ijcai/ZhangLTCL13}.
Meanwhile, building one model per user is not practical.
To this end, this paper presents a novel solution, of which the principle is to model the \retg{} behavior over user groups.
Our system, \sys{}, consists of three key components for extracting user based features, clustering users into groups, and modeling upon each group.
Particularly, we look into the user interest from different perspectives including long-term/short-term interests and explicit/implicit interests.
%, which results deep analyses towards the \retg{} behavior and proper models in the end.
We have evaluated the performance of \sys{} using datasets of real-world social networking applications and a number of query workloads, showcasing its benefits.
\end{abstract}
% behavior and activity in this paper are interchangeable